---
title: "Seurat Label Transfer"
author: 
 - Richard Green
 - Lucas Graybuck
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    df_print: paged
    self_contained: true
params:
  in_h5: NULL
  out_h5: NULL
  out_tbl: NULL
---

```{r Load Libraries}
start_time <- Sys.time()

quiet_library <- function(...) {
  suppressPackageStartupMessages(library(...))
}
quiet_library(Seurat)
quiet_library(SeuratLabeler)
quiet_library(H5weaver)
quiet_library(ggplot2)
quiet_library(uwot)

pals.alphabet.colors <- c("#F0A0FF","#0075DC","#993F00","#4C005C","#191919","#005C31","#2BCE48","#FFCC99","#808080","#94FFB5","#8F7C00","#9DCC00","#C20088","#003380","#FFA405","#FFA8BB","#426600","#FF0010","#5EF1F2","#00998F","#E0FF66","#740AFF","#990000","#FFFF80","#FFE100","#FF5005")

options(stringsAsFactors = FALSE)
```

Declaring start
```{r Declare start}
stm("Starting Seurat PBMC Label Transfer")
```

Argument parsing
```{r Parse arguments}
if(is.null(params$in_h5)) {
  in_h5 <- system.file("testdata/pbmc_1k_v3.h5", 
                         package = "SeuratLabeler")
  out_h5 <- tempfile(fileext = ".h5")
  out_tbl <- tempfile(fileext = ".csv.gz")

} else {
  in_h5 <- params$in_h5
  out_h5 <- params$out_h5
  out_tbl <- params$out_tbl
  
  if(is.null(out_tbl)) {
    out_tbl <- "Not Used"
  }
}

stm(paste0("IN  H5 file         : ", in_h5))
stm(paste0("OUT H5 file         : ", out_h5))
stm(paste0("OUT Cell Label Table: ", out_tbl))
```

```{r Print Arguments}
print(c(
  paste0("IN  H5 file         : ", in_h5),
  paste0("OUT H5 file         : ", out_h5),
  paste0("OUT Cell Label Table: ", out_tbl)
))
```


Load Reference Dataset
```{r Load Reference}
stm("Loading Reference from SeuratLabeler package")
reference_so <- variable_pbmc_10k_v3
```

Load Query Dataset
```{r Load Query}
stm(paste("Loading query from", in_h5))
query_data <- read_h5_dgCMatrix(in_h5,
                                feature_names = "name")
query_so <- CreateSeuratObject(counts = query_data)
```

Normalize Query Dataset
```{r Normalize Query}
stm("Normalizing query dataset")
query_so <- NormalizeData(object = query_so)
```

Find Anchors
```{r Find Transfer Anchors}
stm("Finding Transfer Anchors")
transfer_anchors <- FindTransferAnchors(reference = reference_so,
                                        query = query_so,
                                        normalization.method = "LogNormalize",
                                        features = VariableFeatures(object = reference_so),
                                        reduction = "pcaproject")
```

Transfer labels
```{r Label Transfer}
stm("Transfering labels")
label_predictions <- TransferData(anchorset = transfer_anchors,
                                  refdata = reference_so$celltype)
query_so <- AddMetaData(object = query_so,
                        metadata = label_predictions)
```

Print summary table
```{r Type Counts}
label_summary <- as.data.frame(table(query_so$predicted.id))
names(label_summary) <- c("type_label","n_cells")
label_summary$frac_cells <- label_summary$n_cells / sum(label_summary$n_cells)

knitr::kable(label_summary, caption = "Type Label Summary")
```
Plot score distribution
```{r Score Distribution}
hist(query_so$prediction.score.max,
     breaks = 50,
     xlim = c(0,1))
```

Extract results and plot N Genes vs Scores
```{r Extract Results}
stm("Extracting labeling results")
label_results <- data.frame(cell_barcode = names(query_so$predicted.id),
                            seurat_pbmc_type = as.character(query_so$predicted.id),
                            seurat_pbmc_type_score = query_so$prediction.score.max,
                            n_genes = query_so$nFeature_RNA)
```

Write output table:
```{r Output Table}
if(out_tbl != "Not Used") {
  stm(paste0("Writing results table to ",out_tbl))
  
  write.csv(label_results,
            file = out_tbl,
            row.names = FALSE)
}
```

Plot genes vs scores
```{r Plot Genes vs Scores}
plot_df <- label_results

plot_colors <- data.frame(seurat_pbmc_type = unique(plot_df$seurat_pbmc_type))
plot_colors$seurat_pbmc_type_color <- pals.alphabet.colors[1:nrow(plot_colors)]

ggplot() +
  geom_point(data = plot_df,
             aes(x = seurat_pbmc_type_score,
                 y = n_genes,
                 color = seurat_pbmc_type)) +
  scale_color_manual(breaks = plot_colors$seurat_pbmc_type,
                     values = plot_colors$seurat_pbmc_type_color) +
  scale_y_log10(breaks = c(10, 100, 500, 1000, 1500, 2000, 3000))
```

Run PCA
```{r Run PCA}
stm("Running PCA and Jack Straw")

sample_cells <- function(cl, 
                          sample.size, 
                          seed = 3030) {
  cl.cells <- split(names(cl), cl)
  sampled.cells <- sapply(names(cl.cells), function(x) {
    cells <- cl.cells[[x]]
    if (sample.size >= length(cells)) {
      return(cells)
    }
    to.sample <- pmin(sample.size, length(cells))
    
    set.seed(seed)
    sample(cells, to.sample)
  }, simplify = FALSE)
  sampled.cells <- unlist(sampled.cells)
  return(sampled.cells)
}

query_so <- FindVariableFeatures(object = query_so)
query_so <- ScaleData(object = query_so)
query_so <- RunPCA(object = query_so, npcs = 50)

jackstraw_cells <- sample_cells(query_so$predicted.id, 100)
jackstraw_so <- query_so[,jackstraw_cells]

jackstraw_so <- JackStraw(object = jackstraw_so, 
                          dims = 50, 
                          num.replicate = 50, 
                          verbose = FALSE)
jackstraw_so <- ScoreJackStraw(object = jackstraw_so, 
                               dims = 1:50)

pc_pvals <- jackstraw_so@reductions$pca@jackstraw$overall.p.values[,2]
good_pcs <- pc_pvals < 0.05

pc_embeddings <- query_so@reductions$pca@cell.embeddings
```

Run UMAP
```{r Run UMAP}
stm("Running UMAP projection")

umap_results <- uwot::umap(pc_embeddings[, good_pcs],
                           min_dist = 0.1)
```

Remove objects from memory:
```{r Remove Seurat Objects}
stm("Dropping Seurat Objects from memory")
rm(query_so)
rm(reference_so)
```

Plot clusters on UMAP
```{r Plot UMAP Clusters}
umap_df <- as.data.frame(umap_results)
names(umap_df) <- c("umap1","umap2")

plot_df <- cbind(plot_df, umap_df) 

ggplot() +
  geom_point(data = plot_df,
             aes(x = umap1,
                 y = umap2,
                 color = seurat_pbmc_type)) +
  scale_color_manual(breaks = plot_colors$seurat_pbmc_type,
                     values = plot_colors$seurat_pbmc_type_color)
```

Plot scores on UMAP
```{r Plot UMAP Scores}
ggplot() +
  geom_point(data = plot_df,
             aes(x = umap1,
                 y = umap2,
                 color = seurat_pbmc_type_score)) +
  scale_color_viridis_c()
```

Plot N Features on UMAP
```{r Plot UMAP N features}
ggplot() +
  geom_point(data = plot_df,
             aes(x = umap1,
                 y = umap2,
                 color = n_genes)) +
  scale_color_viridis_c()
```

Add values to HDF5 List.

Labeling results are stored in "/matrix/observations/"  
UMAP results are stored in "/umap/"  
PCA results are stored in "/pca/"

```{r Add results to list object}
stm("Adding results to input data")

h5_list <- rhdf5::h5dump(in_h5)

if(!"observations" %in% names(h5_list$matrix)) {
  h5_list$matrix$observations <- list()
}

h5_list$matrix$observations$seurat_pbmc_type <- label_results$seurat_pbmc_type[match(h5_list$matrix$barcodes, label_results$cell_barcode)]
h5_list$matrix$observations$seurat_pbmc_type_score <- label_results$seurat_pbmc_type_score[match(h5_list$matrix$barcodes, label_results$cell_barcode)]

h5_list$matrix$observations$umap_1 <- plot_df$umap1[match(h5_list$matrix$barcodes, plot_df$cell_barcode)]
h5_list$matrix$observations$umap_2 <- plot_df$umap2[match(h5_list$matrix$barcodes, plot_df$cell_barcode)]

h5_list$pca_dgCMatrix <- as(t(pc_embeddings), "dgCMatrix")
h5_list <- h5_list_convert_from_dgCMatrix(h5_list, target = "pca")
h5_list$pca$features$jackstraw_pvalue <- pc_pvals
```

Write labeled HDF5 file:
```{r Write HDF5 file}
stm(paste0("Writing h5 results to", out_h5))
write_h5_list(h5_list,
              out_h5)
```


```{r Session Info}
sessionInfo()
```

Total time ellapsed
```{r Show Time}
end_time <- Sys.time()
diff_time <- end_time - start_time
time_message <- paste0("Ellapsed Time: ", 
                       round(diff_time, 3),
                       " ", units(diff_time))
print(time_message)
stm(time_message)
stm("HTO count processing complete.")
```
